---
editor: 
  markdown: 
    wrap: 72
---

# Graph Centrality Tests

```{r,message=FALSE}
#| label: libraries
#| message: false

options(repos = c(CRAN = "https://cloud.r-project.org"))
if (!require("remotes")) install.packages("remotes")
pkgs = c(
    "sf",
    "tidyverse",
    "zonebuilder",
    "tmap"
    # "dodgr" # Using the developer version of dodgr
)
remotes::install_cran(pkgs)
sapply(pkgs, require, character.only = TRUE)

require(dodgr)
```

## Loading network

```{r}
#| label: read-sf
 
sf_bogota_2019_full <- st_read(file.path("sf_network","bogota_osm_network.gpkg"),
                          layer = "network_2019")
```

We will use a small subset of the network for this test to speed up the process.

```{r}
#| label: zone4clipping

bog_zone <- zonebuilder::zb_zone("Bogota",n_circles = 2) |>
  st_transform(st_crs(sf_bogota_2019_full))
# zb_view(bog_zone)
```

The following the code will clip the network; also, since we will be using wrong-way-driving, we have to `allow` both directions inone-way links to capture the changes of centrality.


```{r}
#| label: clipping-net

sf_bogota_2019 <- sf_bogota_2019_full[bog_zone,] |> 
  mutate(oneway = if_else(highway %in% c("residential"),"no",oneway))
```


```{r}
#| label: graph-building

graph_bogota <- weight_streetnet(sf_bogota_2019,
                                 left_side = F,
                                 wt_profile_file = "bogota_wp.json",
                                 wt_profile = "motorcar",
                                 keep_cols = c("oneway","lanes","surface","maxspeed"),
                                 turn_penalty = F)

bog_contracted <- graph_bogota |> 
  dodgr_contract_graph()
```

## Assignment of the Weighting Profile 

In order to include the actual speed along each corridor, I will specify 60 different classes
to be applied to each road link. By default, speeds are assigned by road link by 
road type, e.g. any trunk road in Bogotá has a 60 km/h speed. 


```{r}
#| label: custom-wp

# custom_wp <- dodgr::weighting_profiles$weighting_profiles |> filter(name=="motorcar")
custom_penalties <- dodgr::weighting_profiles$penalties |> filter(name=="motorcar")
custom_surfaces <- dodgr::weighting_profiles$surface_speeds |> filter(name=="motorcar")

custom_wp <- data.frame(name = "motorcar",
                        way = paste0("road_",1:60),
                        value = 1,
                        max_speed = as.numeric(1:60))

custom_wp_list <- list(weighting_profiles = custom_wp,
                       surface_speeds = custom_surfaces,
                       pealties = custom_penalties)

wpj <- jsonlite::toJSON (custom_wp_list, pretty = TRUE)
write_lines(wpj,file = "custom_wp_speeds.json")

```

Assigning a speed category based on the speed limit (baseline scenario)

```{r}
#| label: way-by-speed

sf_bogota_2019_cwp <- sf_bogota_2019 |>
  mutate(way_speed = case_when(highway %in%
                                 c("trunk_link","primary_link","primary","trunk")~"road_60",
                                               TRUE ~ "road_30")) |> 
  select(-highway)
```

The following code builds the graph using the new column instead of  `highway`.

```{r}
#| label: graph-building-2

graph_bogota_custom <- weight_streetnet(sf_bogota_2019_cwp,
                                 left_side = F,
                                 wt_profile_file = "custom_wp_speeds.json",
                                 type_col = "way_speed",
                                 wt_profile = "motorcar",
                                 keep_cols = c("way_speed","oneway","lanes","surface","maxspeed"),
                                 turn_penalty = F)

```

### Comparing results of wp method

We calculate the centralities with both weighting profiles

```{r}
graph_baseline_centrality <- graph_bogota |> 
  dodgr_deduplicate_graph() |>
  dodgr_contract_graph() |>
  dodgr_centrality()

graph_cwp_centrality <- graph_bogota_custom |> 
  dodgr_deduplicate_graph() |>
  dodgr_contract_graph() |>
  dodgr_centrality()
```

To validate that the results are identical, we can run the following code

```{r}
identical(graph_baseline_centrality$centrality,graph_cwp_centrality$centrality)
```

## Centrality Calculations

There are four core alternatives for calculating the centrality using `dodgr`. 

```{r}
#| label: cor-check-1

bog_contracted[,
    c("d_weighted", "d", "time_weighted", "time")] |>
  pairs()
```

An initial check on the four alternatives reveal some inconsistencies in
the `time_weighted` column. 

There are `r nrow(bog_contracted[(bog_contracted$d_weighted - bog_contracted$d)<0,])` links in the contracted graph with differences in the weighted distance and the distance. The location of these edges will be inspected after calculating the centrality in @sec-bc-comp.

When multiple paths are available between two nodes in the contracted version of the graph,
`dodgr` takes only the shortest weighted distance, which creates the difference. However,
the resulting weighted time is not corrected and in some cases a `0` is returned.

Given this, we will recalculate the weighted time based on the weighted distance.

```{r}
bog_contracted$time_weighted[bog_contracted$time_weighted==0] <- (3.6*bog_contracted$d_weighted[bog_contracted$time_weighted==0])/case_when(
      bog_contracted$highway[bog_contracted$time_weighted==0] %in% c("trunk_link","primary_link","primary","trunk")~60,
      T~30)
    
```


### `dodgr`

```{r}
#| label: bc-calc

cent_all <- lapply(c("d_weighted", "d", "time_weighted", "time"), \(x) {
  
    df <- bog_contracted |>
      dodgr_centrality(column = x)
  
  tibble(edge_id = df$edge_id, cent = df$centrality) |>
    rename_with(.cols = "cent", .fn = \(.y) {
      paste(.y, x, sep = "_")
    })
  }) |>
  plyr::join_all(by = "edge_id",type = "full")
```

### `igraph`

```{r}
#| label: graph-to-igraph

igraph_bog <- bog_contracted |> 
  dodgr_to_igraph()
```


```{r}
#| label: bc-igraph-calc

cent_all_ig <- cent_all |> 
  left_join(
    tibble(edge_id = bog_contracted$edge_id,
       cent_ig_d = igraph::edge_betweenness(graph = igraph_bog,
                                            weights = bog_contracted$d),
       cent_ig_d_weighted = igraph::edge_betweenness(graph = igraph_bog,
                                                     weights = bog_contracted$d_weighted),
       cent_ig_time = igraph::edge_betweenness(graph = igraph_bog,weights = bog_contracted$time),
       cent_ig_time_weighted = igraph::edge_betweenness(graph = igraph_bog,
                                                        weights = bog_contracted$time_weighted)
       ),
    by = "edge_id")
```


```{r}
#| label: bc-summary

summary(cent_all_ig[,2:9])
```

Check if `igraph` results are identical to the ones obtained with `dodgr`

```{r}
with(cent_all_ig,identical(cent_ig_d_weighted,cent_d_weighted))
with(cent_all_ig,identical(cent_ig_time_weighted,cent_time_weighted))
```

```{r}
#| label: bc-cor-check

pairs(cent_all_ig[,2:9])
```

Joining the results to the `sf` object

```{r}
sf_net <- graph_bogota |>
  dodgr_to_sf() |> 
  left_join(cent_all,by = "edge_id")
```

### A visual comparison (not very useful) {#sec-bc-comp}
 
```{r}
#| label: cor-check

sf_net |>  
  select(cent_d_weighted:cent_time) |> 
  pivot_longer(names_to = "type",names_prefix = "cent_",
               cols = -geometry,
               values_to = "cent") |> 
  ggplot()+
  geom_sf(aes(col = log(cent),linewidth = cent+1))+
  facet_wrap(type~.)+
  scale_color_viridis_c(
    # direction = -1
    )+
  scale_linewidth_continuous(range = c(0.05,0.5),
                             transform = scales::transform_boxcox(p = 2))+
  theme_void()+
  guides(linewidth = "none",)
  
```

Visualising differences

```{r}
sf_net |> 
  transmute(across(cent_d_weighted:cent_time,
                   \(x) (x - cent_d)/cent_d)) |> 
  pivot_longer(names_to = "type",names_prefix = "cent_",
               cols = -geometry,
               values_to = "cent") |> 
  ggplot(aes(cent))+
  geom_histogram()+
  facet_wrap(type~.)+
  scale_x_continuous(limits = c(-1,1))
  
  
```

Inspecting significant changes

```{r}
sf_net |> 
  transmute(across(cent_d_weighted:cent_time,
                   \(x) (x - cent_d)/cent_d)) |> 
  pivot_longer(names_to = "type",names_prefix = "cent_",
               cols = -geometry,
               values_to = "cent") |> 
  mutate(cent = case_when(abs(cent)<0.1~0,
                          cent > 0.6 ~ 1,
                          -cent > 0.6 ~ -1,
                          is.na(cent)~NA)) |> 
  ggplot()+
  geom_sf(aes(col = cent))+
  facet_wrap(type~.)+
  scale_color_distiller(palette = "Spectral", direction = 1)+
  # scale_linewidth_continuous(range = c(0.05,0.5),
  #                            transform = scales::transform_boxcox(p = 2))+
  theme_void()
```

Identifying the links with differences > 0 in d and weighted d

```{r}
sf_net |> 
  mutate(diff_check = (d-d_weighted)==0) |> 
  ggplot(aes(col = diff_check))+
  geom_sf()+
  theme_void()
```

Same for time

```{r}
sf_net |> 
  mutate(diff_check = (time-time_weighted)==0) |> 
  ggplot(aes(col = diff_check))+
  geom_sf()+
  theme_void()
```

*A course assumption is that people decide the route they will use based on the ETA,
also differences in speed limits can only captured if time is used.
So weighted time will be used.*


### `sfnetworks`

```{r}
library(sfnetworks)
```

```{r}

sf_net$time_weighted[sf_net$time_weighted==0] <- (3.6*sf_net$d_weighted[sf_net$time_weighted==0])/case_when(
      sf_net$highway[sf_net$time_weighted==0] %in% c("trunk_link","primary_link","primary","trunk")~60,
      T~30)

bog_sfnet <- sf_net |> 
  as_sfnetwork(directed = TRUE)
```


```{r}
bog_sfnet <- bog_sfnet |>
  activate("edges") |> 
  mutate(cent_sfn_time_weighted = tidygraph::centrality_edge_betweenness(
    weights = time_weighted,
    directed = T))
```

```{r}
bog_sfnet |> 
  st_as_sf("edges") |> 
ggplot()+
  geom_sf(aes(col = cent_time_weighted))+
  theme_void()
```

```{r}
bog_sfnet |> 
  activate("edges") |>
  data.frame() |> 
  select(cent_time_weighted,cent_sfn_time_weighted) |> 
  pairs()
```

```{r}
bog_sfnet |> 
  activate("edges") |>
  data.frame() |> 
  with(identical(cent_time_weighted,cent_sfn_time_weighted))
```


```{r}
p1 = st_geometry(bog_sfnet, "nodes")[2401]
st_crs(p1) = st_crs(bog_sfnet)
p2 = st_geometry(bog_sfnet, "nodes")[1407]
p3 = st_geometry(bog_sfnet, "nodes")[1509]
p4 = st_geometry(bog_sfnet, "nodes")[1840]
st_crs(p3) = st_crs(bog_sfnet)

paths = st_network_paths(bog_sfnet, from = p1, to = c(p2,p3,p4), weights = "time_weighted")
```


```{r}
plot_path = function(node_path) {
  bog_sfnet %>%
    activate("nodes") %>%
    slice(node_path) %>%
    plot(cex = 1.5, lwd = 1.5, add = TRUE)
}

colors = sf.colors(4, categorical = TRUE)

plot(bog_sfnet, col = "grey")
paths %>%
  pull(node_paths) %>%
  walk(plot_path)
plot(c(p1, p2, p3,p4), col = colors, pch = 8, cex = 2, lwd = 2, add = TRUE)
```




